Memo of meeting

### 2018-07-14 paper/Chapters/Chapter2.tex
methodology updated

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

\section{Strategy}

%-----------------------------------
%	SUBSECTION 1
%-----------------------------------
\subsection{Latent class analysis (LCA)}

@@

\subsection{Multilevel latent class analysis (MLCA)}\vspace{-0.3cm}

@@

\subsubsection{Parametric approach}\vspace{-0.3cm}

@@

\begin{itemize}
	\item $P(C_{ij} = t)$ as the probability that the observation is belonging to latent class $t$;
	\item $u_{0j}$ as the random intercept for $j$th individual; 
	\item the random intercept are assumed be normally distributed (i.e. $u_{0j} \sim N(0, \sigma_{u_0}^2)$), the magnitude of the $u_{0j}$ variance ($\sigma_{u_0}^2$) indicates the influence of the individuals (level 2);
	\item $w_j$ is the predictor for individual (level 2), such as age, and/or sex.
\end{itemize}

@@

\begin{equation}
%\begin{aligned}
P(U_{ij1} = s_1, U_{ij2} = s_2, \cdots, U_{ijk} = s_{k}) = \sum_{t=1}^{T}P(C_{ij}=t)\prod_{k=1}^{K}P(U_{ijk} = s_k | C_{ij} = t)
%\end{aligned}
\label{MLCA}
\end{equation}
\vspace{-0.8cm}

@@

\vspace{-0.5cm}
\subsubsection{Non-Parametric approach}\vspace{-0.3cm}

Another approach is a non-parametric MLCA. The $T-1$

>>>>

%-----------------------------------
%	SECTION 1
%-----------------------------------
\section{Latent class analysis (LCA)}\vspace{-0.3cm}

Latent class analysis is a statistical technique that identifies categorical latent (unobserved) class variables on the basis of observed categorical variables \parencite{collins2010latent}. It belongs to the family of latent variable models, and is directly analogous to the factor analysis model. The major difference is that the latent variable in LCA is categorical, not continuous as in factor analysis. The basic assumptions in LCA are independent observations, and local independence, as shown in the fundamental expression of a typical LCA model: \vspace{-0.8cm}

\begin{equation}
P(U_{i1} = s_1, U_{i2} = s_2, \cdots, U_{ik} = s_K) = \sum_{t=t}^{T}P(C_i = t)\prod_{k = 1}^{K}P(U_{ik} = s_k | C_i = t)
\label{LCA}
\end{equation}\vspace{-0.7cm}

\begin{itemize}
	\item $P(U_{i1} = s_1, U_{i2} = s_2, \cdots, U_{ik} = s_k)$ is the probability of observing a particular vector of responses;
	\item $P(C_i = t)$ is the probability that a randomly selected $i$th observation will be in class $t$;
	\item $P(U_{ik} = s_k | C_i = t)$ is the probability of a particular observed response pattern $U_{ik} = s_k$ conditional on membership in latent class $t$.
\end{itemize}

Equation \ref{LCA} indicates that responses for an observation to the measuring variables are independent of one another given its membership in latent class $t$. However, in the NDNS RP data set, the assumption of independent observations is violated. Each individual completed their dietary diary for four consecutive days, their diary recordings were later converted into 4 sequences of categorical responses reflecting the type of carbohydrate consumption at each hour of the day. The 4 observed sequences (observations) are nested in the participants and therefore are not independent. This nested data structure requires multilevel techniques. 
\vspace{-0.3cm}

@@

\section{Multilevel latent class analysis (MLCA)}\vspace{-0.3cm}

@@

\subsection{Parametric approach}\vspace{-0.3cm}

@@

\begin{itemize}
	\item $P(C_{ij} = t)$ as the probability that the randomly selected $i$th observation of $j$th individual is belonging to latent class $t$;
	\item $u_{0j}$ as the random intercept for $j$th individual; 
	\item the random intercept are assumed be normally distributed (i.e. $u_{0j} \sim N(0, \sigma_{u_0}^2)$), the magnitude of the $u_{0j}$ variance ($\sigma_{u_0}^2$) indicates the influence of the individuals (level 2);
	\item $w_j$ is the predictor for individual (level 2), such as age, and/or sex.
\end{itemize}

@@

\begin{equation}
%\begin{aligned}
P(U_{ij1} = s_1, U_{ij2} = s_2, \cdots, U_{ijk} = s_{K}) = \sum_{t=1}^{T}P(C_{ij}=t)\prod_{k=1}^{K}P(U_{ijk} = s_k | C_{ij} = t)
%\end{aligned}
\label{MLCA}
\end{equation}
\vspace{-0.8cm}

@@

\vspace{-0.5cm}
\subsection{Non-Parametric approach}\vspace{-0.3cm}

Another approach is using a non-parametric MLCA, in which separate latent class models are specified for level 1 and level 2. Similar with the parametric MLCA approach, there are $T-1$ random intercepts, where $T$ is the number of level 1 latent classes. However, rather than assuming the random intercepts following a normal distribution, the non-parametric MLCA assumes a multinomial (discrete) distribution of the level 2 latent classes. This approach is less computationally demanding compared with the parametric approach. These level 2 latent
classes reflect differences in the probability of belonging to a specific level 1 latent class, so that clusters (i.e., individuals) that contain observations with similar probabilities for the level 1 latent classes will be grouped together. The non-parametric MLCA model can be defined as follows: \vspace{-0.7cm}

\begin{equation}
P(C_{ij} = t | CB_j = m)  = \frac{\exp(\gamma_{tm})}{\sum_{r=1}^{T}\exp(\gamma_{tm})}
\end{equation}

\begin{itemize}
	\item $CB_j$ is level 2 latent class membership for $j$th individual;
	\item $\gamma_{tm}$ is level 1 and level 2 indicators.
\end{itemize}

According to Finch and French's simulation study \parencite{finch2014multilevel}, non-parametric approach generally resulted in more accurate recovery of the underlying latent structure of the data at both levels and provided better latent class model. Specifically, we are interested in exploring meaningful individual (level 2) latent classes rather than their daily consumption classification. Therefore, non-parametric MLCA was employed 1) to identify latent classes of observations (level 1) based on the subjects' responses to the 4-day food and drink diary and 2) to form distinct latent classes of individuals (level 2) based on the distribution of observation-level latent classes within individuals.
